|

### 2020-07-06 paper.Rmd
re-working abstract

The notion of a "curse of dimensionality" was originally introduced in @BellmanRichard1961 to describe the exponentially growing difficulty of optimization over an increasing number of dimensions. More generally, we can think of it as a problem of sampling large parameter spaces. As the number of dimensions increases, the volume of the parameter space grows exponentially, and the density of points decreases. As a result observations will be far from the center of the distribution, and in the extreme case of high-dimension low-sample size date observations are on the vertices of a simplex [@doi:10.1111/j.1467-9868.2005.00510.x]. This causes problems in different domains. An important example in statistics is cluster analysis. Since the distance between any two points in a distribution becomes approximately constant, new methods are needed to cluster points in a high-dimensional space (see e.g. @Steinbach2004).

A more detailed discussion of the curse of dimensionality was presented in @Donoho00, which also highlighted the opposite effect, a blessing of dimensionality. Indeed we can sometimes exploit large parameter spaces to our advantage. One example is the use of regularization methods for variable selection. These methods penalize model complexity in the optimization, as for example in ridge regression and Lasso. The penalty results in shrinking (some of) the parameter estimates towards zero. In ridge regression we obtain a global shrinkage, while Lasso can be thought of as variable selection and will shrink some of the estimates to zero. XXX mention concentration of measures?

>>>>

The notion of a "curse of dimensionality" was originally introduced in @BellmanRichard1961 to describe the exponential growth in difficulty for optimization as number of dimensions increases. This affects sampling in high-dimensions. The volume of the space grows exponentially, with dimension, and the density of points decreases proportionally (??). The effect is that most points will be far from the sample mean, on the edge of the sample space. @doi:10.1111/j.1467-9868.2005.00510.x have shown that in the extreme case of high-dimension, low-sample size data, observations are on the vertices of a simplex.

<!-- This causes problems in different domains. An important example in statistics is cluster analysis. Since the distance between any two points in a distribution becomes approximately constant, new methods are needed to cluster points in a high-dimensional space (see e.g. @Steinbach2004).-->

@Donoho00, conversely, considers the curse of dimensionality as a blessing, because the sparsity can be leveraged for computational efficiency. This is used in regularization methods, like lasso, to penalize model complexity. The penalty results in shrinking (some of) the parameter estimates towards zero.<!-- In ridge regression we obtain a global shrinkage, while Lasso can be thought of as variable selection and will shrink some of the estimates to zero. XXX mention concentration of measures?-->
